\documentclass[./main.tex]{subfiles}

\begin{document}


\section{Kernels and Ideals}
When we studied groups, we studied structure-preserving maps between groups. Now
it is time to study structure-preserving maps between rings. 

\begin{definition}[Ring homomorphisms]
\label{def:ring-homomorphisms}
Let $R, S$ be a rings and $\varphi: R \to S$. Then $\varphi$ is a \textbf{ring
homomorphism} if $\varphi$ is a group homomorphism: $\varphi(a+b) = \varphi(a) +
\varphi(b)$ for $a,b \in R$ and $\varphi$ is multiplicative: $\varphi(ab) =
\varphi(a) \varphi(b)$.
\end{definition}

Some authors (e.g. \cite{Jacobson_2009}) generally require the following axiom:
If additionally both $R, S$ have a unity, then $\varphi(1) = 1$. This
stipulation is made in order to ensure that units are mapped onto units. 

We also make the definition of a 
\begin{definition}[Kernel]
\label{def:ring-homomorphism-kernel}
The \textbf{kernel} of a ring homomorphism $\varphi$ is the set 
\[
    \ker \varphi = \set{r \in R: \varphi(r) = 0}.
\]
\end{definition}

% TODO: Include D&F ch 7.3 exercise 17 (p263 in pdf)

\begin{example}
    For $n > 0 \in \bZ$, there is a canonical group homomorphism $\bZ \to \bZ_n$
    given by $m \mapsto m \pmod n$. Since modular arithmetic is multiplicative,
    this map is in fact a ring homomorphism. 
\end{example}

\begin{example}
    If $n \in \bZ$, the map $\varphi_n: \bZ \to \bZ$ defined by $m \mapsto mx$
    is not a ring homomorphism unless $n^2 = n$. However $\varphi_n$ is always a
    group homomorphism on the additive group of $\bZ$. This example emphasizes
    the importance of checking the multiplicativity of a map.
\end{example}

\begin{example}[Evaluation homomorphism]
    Let $R$ be a commutative ring with unity, let $r \in R$, and let
    $\operatorname{eval}_r: R[x] \to R$ denote the mapping $f(x) \mapsto f(r)$.
    This is called the \emph{evaluation homomorphism} because it simply
    evaluates the polynomial $f$ at $r$. We leave it to the reader to check that
    this is indeed a homomorphism. 
\end{example}
% We digress to comment on the evaluation map. On one hand $R[x]$ is the set of
% formal polynomials in $x$ -- sequences of elements in $R$. Yet, we can connect
% them with polynomial functions $R \to R$ using the evaluation homomorphism.

Let us now look at a few properties of ring homomorphisms. 

\begin{proposition}
    Let $R, S$ be rings and $\varphi: R \to S$ be a homomorphism. Then,
    $\varphi[R]$ is a subring of $S$, and $\ker \varphi$ is a subring of $R$. 
\end{proposition}
\begin{proof}
    Routine.
\end{proof}
\begin{exercise}
    Prove the above proposition.
\end{exercise}

The fact that $\ker \varphi$ is a subring of $R$ is only part of the story
though. If $r \in R$, and $a \in \ker \varphi$, then $\varphi(ra) = \varphi(r)
\varphi(a) = 0$, so $ra \in \ker \varphi$ too; similary for $ar$. Such subrings
are given a special name:

\begin{definition}[Ideal]
\label{def:ideal}
    A(n) (two-sided) \textbf{ideal} $I$ of a ring $R$ is a subring of $R$ such
    that for all $i \in I$ and all $r \in R$, both $ri$ and $ir$ are in $R$.
\end{definition}

Note if we only assume that $ri \in R$ then $I$ is a left ideal, and if only $ir
\in R$ we call $I$ a right ideal. So a two-sided ideal is both a left and right
ideal. Some authors do define left and right ideals, but their utility is rather
limited and so we have chosen to omit them. Nevertheless, the concept of
one-sided ideals will be developed in the exercises for the interested reader.
Henceforth, whenever we say ideal, we shall mean two-sided ideal as in
\cref{def:ideal}.

It would be rather silly to give a name to a concept that doesn't have utility.
It turns out the utility of ideals is much like normal subgroups: they help us
form quotient rings. 

\begin{proposition}[Existence of quotient rings]
\label{prop:existence-of-quotient-rings}
    Let $I \subseteq R$ be a subring. Then $I$ is an ideal of $R$ if and only if
    the operation $(a + I)(b + I) := ab + I$ is well-defined. if this operation
    is well-defined, $R/I$ is a quotient ring with multiplication given by said
    operation.
\end{proposition}


\begin{proof}
    Recall that $R/I$ is denoted to be the set of left cosets of $I$. We shall
    denote the left coset $\set{a + i: i \in I}$ by $a + I$. Clearly $R/I$ is
    already a quotient group since $R$ is abelian. So all that's left to check is
    the ring operation. We leave checking that as an exercise.
\end{proof}

\begin{exercise}
    Complete the proof of \cref{prop:existence-of-quotient-rings}.
\end{exercise}

Just like how the study of group homomorphisms is intimately related to the
study of normal subgroups, so is the study of ring homomorphisms with the study
of ideals.

\begin{proposition}
\label{prop:all-ideals-are-kernels}
    Every ideal is the kernel of some ring homomorphism
\end{proposition}
\begin{proof}
    Consider $r \mapsto r + I$.
\end{proof}

Any time we see quotients, the first isomorphism theorem should come to mind.
The reader has probably already formulated the first isomorphism theorem for
rings, but we shall state it anyway.

\begin{theorem}[First Isomorphism Theorem (Rings)]
\label{thm:first-isomorphism-thm-rings}
Let $R, S$ be rings and let $\varphi: R \to S$ be a ring homomorphism. Then, the
map $\tilde \varphi: R/\ker \varphi \to \varphi[S]$ given by $\tilde \varphi(r
\ker \varphi) = \varphi(r)$ is an isomorphism. Diagramatically,

\[\begin{tikzcd}
	R & {\varphi[R]} \\
	{R / \ker \varphi}
	\arrow["\varphi", from=1-1, to=1-2]
	\arrow["\pi"', from=1-1, to=2-1]
	\arrow["\tilde\varphi"', dashed, from=2-1, to=1-2]
\end{tikzcd}\]

\end{theorem}

\begin{proof}
    Exercise.
\end{proof}
\begin{exercise}
    Prove \cref{thm:first-isomorphism-thm-rings}. 
\end{exercise}

Now that we are sufficiently convinced of the utility of ideals, let's see some
examples. In the following examples, notice how we construct homomorphisms to
avoid directly arguing that a certain set is an ideal.

\begin{example}
    The kernel of any homomorphism is an ideal. Note that
    \cref{prop:all-ideals-are-kernels} shows that these are precisely all the
    ideals. 
\end{example}

\begin{example}
    The trivial ideal $\set{0}$ and the whole ring $R$ are both ideals. This can
    be done by definition, or by observing that the whole ring $R$ is the kernel
    of the trivial homomorphism, and the trivial ideal is the kernel of the
    identity homomorphism $R \to R$.
\end{example}

\begin{example}[The ideals of the integers]
    Let's analyze what the ideals of $\bZ$ can be -- this will completely
    characterize \emph{all} homomorphisms out of $\bZ$. Of course, we should
    tackle the easier task first, finding exmaples of ideals of $\bZ$. Since the
    only subgroups of $\bZ$ are cyclic of the form $\gen{a}$, this severely
    limits candidates for ideals. We claim these are all in fact ideals. To see
    this, for each $n > 0$, let $\phi_n$ be the map $m \mapsto n \mod n$. It's
    not hard to check that $\phi_n$ is a ring homomorphism, and that $\ker
    \phi_n$ is precisely $\gen n$. So all the ideals of $\bZ$ are of the form
    $\gen n$.
    
    An alternative argument for this classification is possible by directly
    using the definition of an ideal and the well-ordering principle; we leave
    this to the reader.
\end{example}

\begin{example}[Kernel of the evaluation homomorphism]
    We have previously discussed the evaluation homomorphism in a previous
    example. Let $\varphi$ denote the evaluation homomorphism from $R[x] \to R$
    given by $p(x) \mapsto p(0)$. What is $\ker \varphi$? Well, $\varphi(p) = 0$
    if and only if $p(0) = 0$. So the ideal is the set of all polynomials with
    constant term 0. We will soon introduce notation that will allow us to write
    this in a cleaner manner.
\end{example}


\section{Construction of ideals}
We've seen quite a few examples of ideals, but how do we construct ideals? One
way is to define ring homomorphisms, and calculate the kernel. However,
sometimes this can be annoying. Hence, we shall now introduce some operations
which allow us to create ideals. For this section, $R$ will always be a ring
with unity.


\begin{definition}[Ideal generated by a set]
\label{def:generation-of-ideals}
Let $R$ be a ring with unity, and let $A \subseteq R$ be \emph{any} subset. Then the
\textbf{(two-sided) ideal generated by $A$}, denoted $(A)$, is the smallest (two-sided) ideal of $R$
that contains $A$. Equivalently, it is the intersection of all the (two-sided) ideals of $R$
that contain $A$.

If $A$ is finite, i.e. $A = \set{a_1, \dots, a_n}$, then we can also write
$(a_1, \dots, a_n)$. If $A$ is a single element, i.e. $A = \set{a}$, then we
write $(a)$.
\end{definition}
If $A$ is finite, then the ideal generated by $A$ is said to be \textbf{finitely
generated}. If $A$ is a singleton, the ideal generated by $A$ is called a
\textbf{principal ideal}.

Since the whole ring $R$ is an ideal, the set of ideals containing $A$ is
nonempty, and hence the intersection of this set is a well-defined concept. One
can also equivalently define the left or right ideal generated by $A$. We shall
shortly see that these are related to the next concept we shall introduce:


\begin{definition}
    Let $R$ be a ring and $A \subseteq R$ be \emph{any} set.
    
    Let $RA$ denote the set of all finite $R$-linear combinations of 
    elements in $A$, i.e. 
    \[
        RA = \Set{\sum_{\text{finite}} r_i a_i : r_i \in R, a_i \in A}.
    \]
\end{definition}
We can define $AR$ in a similar manner, and $RAR$ would be the set of all finite
sums of elements of the form $r_i a_i r'_i$.
It turns out 



\end{document}